Designing and modeling a browser-based distributed evolutionary computation systems

### 2015-03-31 modeling-volunteer-computing-gecco.tex
Answering @fergunet reviews

% Había metido en el .bib el artículo de detectar vulnerabilidades 
% en EAs distribuidos de Juanlu, pero creo que al final no tiene sentido 
% hablar en este paper de él (Muszynski14vulnerabilities)
%Lo voy a meter
In fact, this relationship of trust is in many cases the only
mechanism preventing the malicious use of the system by creating bogus
solutions or simply attempting to exploit vulnerabilities by injecting
code. These problems have been analyzed in
\cite{Muszynski14vulnerabilities} but in principle our working
hypothesis will be that the clients using the browser are not
manipulating the code make the experiment finish before it should,
which is possibly the only way in which the experiment would be
affected.

@@

The rest of the paper is organized as follows: next we will present
the state of the art in volunteer computing and its modelization. We 
%FERGU: aunque hay artículos con el término "modelization", no sé si el correcto será "modelling"
will proceed to describe the experimental setup in Section
\ref{sec:exp}, some preliminary results in Section \ref{sec:res} and
finally we will present our conclusions in Section \ref{sec:conc}
along with future lines of work.

@@

Using the web as a resource for distributed, or plainly user based,
evolutionary computation has a long history since JavaScript was
introduced as a browser-based language in 1997 and even before, when
other procedures, including Flash animations, VBScript, ActiveX or Java applets were
used. Java was pointed out in \cite{soares1998get} as a ``language for
internet'', providing some advantages such as multi-architecture compatibility or 
security mechanisms.
%\begin{quote}
%Java can bring some important advantages... solves architectural
%heterogeneity [...] easy to install and [...] built-in security
%mechanisms
%\end{quote} FERGU: quito esta cita y lo pongo sin el quote
In that work, Soares et al. describe JET, a system that supports
the execution of parallel applications over the Internet. This system has 
a comprehensive statistics support, allowing its use for science, 
unlike the other systems compared in that paper.

@@

There have been relatively few efforts to analyze what is the
performance that can be obtained from these volunteer computing
effort. There was some attempt initially to dodge the issue by making
the algorithm adaptive to the kind of resources allotted to it
\cite{milani2004online}, which is actually not such a big problem in
algorithms such as the EA that can easily be
parallelized via population splitting or farming out evaluations to all
the nodes available. Lately, several approaches have focused on the
fault-tolerance of volunteer algorithms
\cite{gonzalez2010characterizing} which can, of course, be studied in
a more general distributed computing context
\cite{nogueras2015studying} or including it in a more general study of
performance of the EA itself %Con tu permiso lo paso a acrónimos para que no sea tan largo el párrafo. FERGU
\cite{DBLP:journals/gpem/LaredoBGVAGF14}. But the raw material of
volunteer computing, number of users and the time spent in the
computation in browser-based volunteer computing experiments, have only been studied in a limited way in
\cite{DBLP:journals/gpem/LaredoBGVAGF14} on the basis of a single
run. Studies using volunteer computing platforms such as SETI@home
\cite{javadi2009mining} found out that the Weibull, log-normal and Gamma distribution
modeled quite well the availability of resources in several clusters
of that framework; the shape of those distributions is a skewed bell
with more resources in the {\em low} areas than in the high areas:
there are many users that give a small amount of cycles, while there
are just a few that give many cycles. This is in concordance with the
results obtained in \cite{agajaj}.

As far as we know, this paper presents the only experiment that uses computational
resources that are as dissimilar as smartphones and powerful laptops
or desktop computers in a research center. The methodology used to
gather resources and the algorithms used will be described next.

%FERGU: volver a indicar que además se dan resultados de uso, para que quede clara la contribución?

\section{Methodology}
\label{sec:exp}

In order to test the volunteer computing environment, a presentation
describing a low cost volunteer computing environment was
created. This presentation was actually delivered in several
conferences (such as IDC 2014). During the %FERGU: TODO JJ, no te olvides de decirme esto. Es la IDC 2014?
conference, it was revealed to the persons attending it that they were participating
in the experiment. The same procedure was used when trying to gather
users in social networks: a description of the container (the
presentation) and disclosure of its purpose. The upper right corner of
the presentation shows the progress of the evolution. It
stops when the solution has been reached.

@@

The server was also written using node.js and the {\tt express.js}
module and has three REST \cite{Castillo12REST} %FERGU: cito a este paper que es un capítulo de libro en vez de Arxiv
functions, that is, three functions that can be accessed from any
client, including the browser, via a request encoded in an URL and a
standard HTTP command. These functions are\begin{itemize}
\item {\tt GET random} returns a random, non-evaluated individual,
  from the pool. The pool is just an array containing chromosomes,
  that is initialized at the beginning of the experiment.
\item {\tt PUT one} puts a single individual along with its fitness in the pool. In REST
  conventions, {\tt PUT} is used to create a resource, that is why it
  is used, in this case, to add a new individual to the pool.
\end{itemize}

@@

As indicated in the introduction, every part of the experiment and
data gathered have been released as free software in the GitHub
repository \url{https://github.com/JJ/idc-keynote}. %FERGU: JJ, es este el repo correcto, no?
These results
will be analyzed in the next section.

@@

At any rate, this exponential trend in the first posts breaks down
after a number of IPs. That could be due to several possible factors:
persons checking the web for just an instant, late comers, or very
slow computers that are not able to lend more cycles to the
simulation. The first IP in both cases contributes on the order of a
hundred PUTs, that is, several thousands generations and several
hundred thousands evaluations, which, once again, is quite remarkable
for such a simple experiment announced by an ephemeral tweet. It
should be noted that, since the speed in the browser virtual machines
(Chrome, mainly, but also Firefox) is as fast as the node.js version,
a parallel version does not add overhead to the single-browser
version, which, as mentioned above, can run in an independent
way: a second browser always adds to the first, even if the total
number of evaluations and the time needed will not speed up in the
same way. However, even taking into account the stochastic nature of a
single evolutionary algorithm, what we are going to find is a high
dispersion of the times needed to reach a solution, which are plotted
in Figure \ref{fig:t} against the number of IPs. There is a wide
range of durations, although there is no clear decrease of the time
needed as the number of IPs used increase. However, we will have to
investigate further the speedups achieved and if these are %FERGU: speedup va junto
algorithmic or due to the particular asynchronous and web-based
implementation. Figure \ref{fig:intervals} might help understand the
kind of raw material we are using: It plots a histogram of the time
needed by all experiments to process 100 generations, 25600
evaluations and the associated evolutionary geneticry. The vast %FERGU: geneticry?
majority of browsers processes it in less than 3 seconds (logarithm( 
time ) < 0.5, which is equivalent to 3.16). However, it is interesting
to note that there are some intervals that go up to the thousands of
seconds, for which we cannot really offer an explanation, other than
the user leaving the page and returning after a while. However, 75\%
of the cases take less than 4 seconds, which is thus a statistical
measure of the kind of performance we can expect from clients for this
type of problem. We can also conclude that 75\%, or a majority, will
be in the fastest tier, although there will be a small amount of them
that will be quite slow and will probably be impossible to accommodate
in the distributed computing experiment.

>>>>

% Hab√≠a metido en el .bib el art√≠culo de detectar vulnerabilidades 
% en EAs distribuidos de Juanlu, pero creo que al final no tiene sentido 
% hablar en este paper de √©l (Muszynski14vulnerabilities)
%Lo voy a meter
In fact, this relationship of trust is in many cases the only
mechanism preventing the malicious use of the system by creating bogus
solutions or simply attempting to exploit vulnerabilities by injecting
code. These problems have been analyzed in
\cite{Muszynski14vulnerabilities} but in principle our working
hypothesis will be that the clients using the browser are not
manipulating the code make the experiment finish before it should,
which is possibly the only way in which the experiment would be
affected.

@@

The rest of the paper is organized as follows: next we will present
the state of the art in volunteer computing and its modelling. We 
%FERGU: aunque hay art√≠culos con el t√©rmino "modelization", no s√©
%si el correcto ser√° "modelling"
% OK- JJ
will proceed to describe the experimental setup in Section
\ref{sec:exp}, some preliminary results in Section \ref{sec:res} and
finally we will present our conclusions in Section \ref{sec:conc}
along with future lines of work.

@@

Using the web as a resource for distributed, or plainly user based,
evolutionary computation has a long history since JavaScript was
introduced as a browser-based language in 1997 and even before, when
other procedures, including Flash animations, VBScript, ActiveX or Java applets were
used. Java was pointed out in \cite{soares1998get} as a ``language for
internet'', providing some advantages such as multi-architecture compatibility or 
security mechanisms.
%\begin{quote}
%Java can bring some important advantages... solves architectural
%heterogeneity [...] easy to install and [...] built-in security
%mechanisms
%\end{quote} FERGU: quito esta cita y lo pongo sin el quote
% OK, pero øpor? - JJ
In that work, Soares et al. describe JET, a system that supports
the execution of parallel applications over the Internet. This system has 
a comprehensive statistics support, allowing its use for science, 
unlike the other systems compared in that paper.

@@

There have been relatively few efforts to analyze what is the
performance that can be obtained from these volunteer computing
effort. There was some attempt initially to dodge the issue by making
the algorithm adaptive to the kind of resources allotted to it
\cite{milani2004online}, which is actually not such a big problem in
algorithms such as the EA that can easily be
parallelized via population splitting or farming out evaluations to all
the nodes available. Lately, several approaches have focused on the
fault-tolerance of volunteer algorithms
\cite{gonzalez2010characterizing} which can, of course, be studied in
a more general distributed computing context
\cite{nogueras2015studying} or including it in a more general study of
performance of the EA itself %Con tu permiso lo paso a acr√≥nimos para
                             %que no sea tan largo el p√°rrafo. FERGU
%OK - JJ
\cite{DBLP:journals/gpem/LaredoBGVAGF14}. But the raw material of
volunteer computing, number of users and the time spent in the
computation in browser-based volunteer computing experiments, have only been studied in a limited way in
\cite{DBLP:journals/gpem/LaredoBGVAGF14} on the basis of a single
run. Studies using volunteer computing platforms such as SETI@home
\cite{javadi2009mining} found out that the Weibull, log-normal and Gamma distribution
modeled quite well the availability of resources in several clusters
of that framework; the shape of those distributions is a skewed bell
with more resources in the {\em low} areas than in the high areas:
there are many users that give a small amount of cycles, while there
are just a few that give many cycles. This is in concordance with the
results obtained in \cite{agajaj}.

As far as we know, this paper presents the only experiment that uses computational
resources that are as dissimilar as smartphones and powerful laptops
or desktop computers in a research center. The methodology used to
gather resources and the algorithms used will be described next, as
well as the results obtained with the setup.

%FERGU: volver a indicar que adem√°s se dan resultados de uso, para
%que quede clara la contribuci√≥n?
%JJ: Hecho.

\section{Materials: running an island model in the browser}
\label{sec:exp}

In order to test the volunteer computing environment, a presentation
describing a low cost volunteer computing environment was
created. This presentation was actually delivered in several
conferences (such as IDC 2014). During the %FERGU: TODO JJ, no te
                                %olvides de decirme esto. Es la IDC
                                %2014?
% Y el PPSN. Pero eso lo voy a dejar para el trabajo anterior, voy a
% intentar hacer experimentos nuevos. - JJ
conference, it was revealed to the persons attending it that they were participating
in the experiment. The same procedure was used when trying to gather
users in social networks: a description of the container (the
presentation) and disclosure of its purpose. The upper right corner of
the presentation shows the progress of the evolution. It
stops when the solution has been reached.

@@

The server was also written using node.js and the {\tt express.js}
module and has three REST \cite{Castillo12REST} %FERGU: cito a este
                                %paper que es un cap√≠tulo de libro en
                                %vez de Arxiv
% OK - JJ
functions, that is, three functions that can be accessed from any
client, including the browser, via a request encoded in an URL and a
standard HTTP command. These functions are\begin{itemize}
\item {\tt GET random} returns a random, non-evaluated individual,
  from the pool. The pool is just an array containing chromosomes,
  that is initialized at the beginning of the experiment.
\item {\tt PUT one} puts a single individual along with its fitness in the pool. In REST
  conventions, {\tt PUT} is used to create a resource, that is why it
  is used, in this case, to add a new individual to the pool.
\end{itemize}

@@

As indicated in the introduction, every part of the experiment and
data gathered have been released as free software in the GitHub
repository \url{https://github.com/JJ/idc-keynote}. %FERGU: JJ, es
                                %este el repo correcto, no?
% Todo esto lo voy a cambiar, no lo mires mucho m·s que por
% metodologÌa. Ahora uso http://github.com/JJ/splash-volunteer
These results
will be analyzed in the next section.

@@

At any rate, this exponential trend in the first posts breaks down
after a number of IPs. That could be due to several possible factors:
persons checking the web for just an instant, late comers, or very
slow computers that are not able to lend more cycles to the
simulation. The first IP in both cases contributes on the order of a
hundred PUTs, that is, several thousands generations and several
hundred thousands evaluations, which, once again, is quite remarkable
for such a simple experiment announced by an ephemeral tweet. It
should be noted that, since the speed in the browser virtual machines
(Chrome, mainly, but also Firefox) is as fast as the node.js version,
a parallel version does not add overhead to the single-browser
version, which, as mentioned above, can run in an independent
way: a second browser always adds to the first, even if the total
number of evaluations and the time needed will not speed up in the
same way. However, even taking into account the stochastic nature of a
single evolutionary algorithm, what we are going to find is a high
dispersion of the times needed to reach a solution, which are plotted
in Figure \ref{fig:t} against the number of IPs. There is a wide
range of durations, although there is no clear decrease of the time
needed as the number of IPs used increase. However, we will have to
investigate further the speedups achieved and if these are %FERGU:
                                %speedup va junto
%si son correcciones de estas no hace falta que las comentes, hazlas y
%ya est· - JJ
algorithmic or due to the particular asynchronous and web-based
implementation. Figure \ref{fig:intervals} might help understand the
kind of raw material we are using: It plots a histogram of the time
needed by all experiments to process 100 generations, 25600
evaluations and the associated evolutionary geneticry. The vast
%FERGU: geneticry?
%JJ un palabro inventao 
majority of browsers processes it in less than 3 seconds (logarithm( 
time ) < 0.5, which is equivalent to 3.16). However, it is interesting
to note that there are some intervals that go up to the thousands of
seconds, for which we cannot really offer an explanation, other than
the user leaving the page and returning after a while. However, 75\%
of the cases take less than 4 seconds, which is thus a statistical
measure of the kind of performance we can expect from clients for this
type of problem. We can also conclude that 75\%, or a majority, will
be in the fastest tier, although there will be a small amount of them
that will be quite slow and will probably be impossible to accommodate
in the distributed computing experiment.
